---
title: "MidTerm_Response_Jack_Tan"
author: "JACKTANSNAKE"
date: "2021/2/16"
output: html_document
---

```{r}
library(survival)
library(tidyverse)
```

```{r}
AppPopularity = read.csv( 'https://www.macalester.edu/~addona/AppPopularity.csv' )
```

## 1.

### a.

```{r}
KM <- survfit(Surv(DaysSurvived, Status) ~ Type , data = AppPopularity, conf.type = "plain")
summary(KM)
```

Free:

26, [18, 45]

Paid:

75, [49, 83]

Yes we do have enough evidence that there is truly a difference in these 80th quantiles in that the two 95% CIs does not intersect.

### b.

No, they are not symmetric.

### c. 

##### i.

```{r}
survdiff(Surv(DaysSurvived, Status) ~ AgeRate, data = AppPopularity)
```

Hypothesis:

$H_0$: $S_{AgeRate=04}(t) = S_{AgeRate=09+}(t) = S_{AgeRate=12+}(t) = S_{AgeRate=17+}(t)$ for all t.

$H_a$: $S_{AgeRate=04}(t) \neq S_{AgeRate=09+}(t) \neq S_{AgeRate=12+}(t) \neq S_{AgeRate=17+}(t)$ for some t.

using the Hanzel-Cox method:

test statistics:

```{r}
0.740 + 0.307 + 1.310 + 2.159
```

p-value:

```{r}
1 - pchisq(4.516, df = 3)
```

Conclusion:

Since the p-value is greater than 0.05, we do not have enough evidence to say that the survival of different `AgeRate` are no identical.

##### ii.

Question a. was asking about the difference of survival at a certain time point, whereas Question c. ii. is asking about if the entire survival curves are identical or not.

### d.

```{r}
AppPopularity %>%
  filter(Status == 1) %>%
  distinct(DaysSurvived)
```

242 tables.

### e.

The lowest one will be the curve having the smallest $\frac{(O-E)^2}{E}$ and the highest one will be the curve having the largest $\frac{(O-E)^2}{E}$.

### f.

```{r}
m = survreg(Surv(DaysSurvived, Status) ~ AvgRate, dist ='lognormal', data = AppPopularity)

summary(m)
```

```{r}
1 - plnorm(50, 1.3860 + 0.1842 * 3.5, 1.8)
```

About 14.8% chance that an app with 3.5 rating would survive longer than 50 days.

### g.

```{r}
qlnorm(1 - 0.14797, 1.3860 + 0.1842 * 4.5, 1.8)
```

Around 60 days.

### h.

$$
\frac{60.11282}{50} = exp(0.1842)
$$

## 2.

```{r}
life_table <- data.frame(Intervals = c("0-1", "1-2", "2-3", "3-4", "4-5", "5-6"),Interval_Failure_rate = c(0,0,0,0,0,NaN), Interval_Survival_rate = c(0,0,0,0,0,NaN), Survival_at_left_end_pt = c(1,0.8,0.6,0.4,0.3,0.05))
```

```{r}
for (i in 1:5){
  life_table[i, 3] <- life_table[i+1, 4]/ life_table[i, 4]
}

for (i in 1:5){
  life_table[i, 2] <- 1 - life_table[i, 3]
}

life_table
```

Since the interval length is 1, `Interval Failure Rate` is the same as the estimate of hazard in each interval. For interval `0-1`, the hazard estimate is 0.2; for interval `1-2`, the hazard estimate is 0.25; for interval `2-3`, the hazard estimate is 0.333333; for interval `3-4`, the hazard estimate is 0.25; for interval `4-5`, the hazard estimate is 0.8333333.

## 3.

### a.

False

### b.

True

### c.

False

### d.

False

### e.

True

### f.

True

## 4.

[0, 25):   Group 2's hazard is higher 

[25, 50):  Group 1's hazard is higher

[50, 75):  Group 2's hazard is higher

[75, 100): Hard to tell which group has a higher hazard 

## 5.

### a.

right-censored data. Because we only know the lower bound of the time of weathering, but do not know the exact weathering time.

### b.

##### i.

P(X < 44) = $\frac{10}{48} = \frac{5}{24}$

##### ii.

\begin{align*}
P(X>26.8) &= \frac{48}{48} = 1\\
P(X>29.6) &= 1 \times \frac{47}{47} = 1\\
P(X>33.4) &= 1 \times \frac{46}{46} = 1\\
P(X>35.0) &= 1 \times \frac{45}{45} = 1\\
P(X>36.3) &= 1 \times \frac{43}{44} = \frac{43}{44}\\
P(X>40.0) &= \frac{43}{44} \times \frac{43}{43} = \frac{43}{44}\\
P(X>41.7) &= \frac{43}{44} \times \frac{41}{42} = \frac{1763}{1848}\\
P(X>41.9) &= \frac{1763}{1848} \times \frac{41}{41} = \frac{1763}{1848}\\
P(X>42.5) &= \frac{1763}{1848} \times \frac{40}{40} = \frac{1763}{1848}\\
P(X>44.0) &= P(X>43.9) = \frac{1763}{1848} \times \frac{38}{39} = \frac{66994}{72072} 
\end{align*}

### c.

```{r}
sweibull <- function(x) 1 - pweibull(x, shape = 1 / 0.06131651, scale = exp(4.025853))
integrate(sweibull, lower = 0, upper = Inf)
```

The mean of `Strength` is 54.24244.

### d.

Because for damaged, those that are not damaged(`Damaged` = 0) are exact data points and the damaged are actually right-censored data points. In this sense, we have to reverse the 0's and 1's in the `Damaged` variable so that it will fit the survival package's requirement.

## 6.

### a.

We assume that the censored observations are at risk for only half of the interval in which they were recorded on average.

### b.

When our data show a non-monotonic trend.

### c.

$$
f(t) = \lambda e^{-\lambda t}
$$

$$
F(t) = \int_{0}^{t} \lambda e^{-\lambda t} = 1 - e^{-\lambda t}
$$

$$
S(t) = 1 - F(t) = e^{-\lambda t}
$$

$$
h(t) = \frac{f(t)}{S(t)} = \frac{\lambda e^{-\lambda t}}{e^{-\lambda t}} = \lambda = \frac{1}{\beta}
$$

$$
\therefore \beta = \frac{1}{h(t)}
$$

$$
log(\beta) = log(\frac{1}{h(t)}) = -log(h(t)) = c_0+c_1x_1+c_2x_2+\cdots+c_kx_k
$$



## 7. 

### a.

Finding the Area under the Curve of a survival curve.

### b.

```{r}
sweibull <- function(x) 1 - pweibull(x, shape = 2, scale = 3)
integrate(sweibull, lower = 0, upper = Inf)
```

The mean survival time is 2.658681.

### c.

$$
S_5(x) = P(X > x|X > 5) = \frac{P(X > x \land X > 5)}{P(X>5)}
$$
Since we have x > 5, we could then have:

$$
S_5(x) = \frac{P(X > x \land X > 5)}{P(X>5)} = \frac{P(X > x)}{P(X>5)} = \frac{S(x)}{S(5)}
$$

For any x > 5, $S_5(x)$ is the original S(x) divided by S(5).

### d.

$$
S_5(x) = P(X > x|X > 5) = \frac{P(X > x \land X > 5)}{P(X>5)}
$$

Since we have $x \leq 5$, we then have:

$$
S_5(x) = \frac{P(X > x \land X > 5)}{P(X>5)} = \frac{P(X > 5)}{P(X>5)} = 1
$$

Thus, for any $x \leq 5$, $S_5(x)$ is equal to 1.

### e. 

```{r}
s5weibull <- function(x) (1 - pweibull(x, shape = 2, scale = 3))/(1 - pweibull(5, shape = 2, scale = 3))
integrate(s5weibull, lower = 0, upper = Inf)
```